---
title: "DDSAnalyticsAttrition"
author: "BalasubramaniamDakshinamoorthi"
date: "12/3/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyr") ## install
library(tidyr)          ## load
install.packages("dplyr") ## install
library(dplyr)          ## load
install.packages("stringr") ## install
install.packages("reshape2") ## install
library(tidyverse)
library(reshape2)
library(ggplot2)
library(stringr)
install.packages("mlbench")
library(mlbench)
install.packages("caret")
library(caret)
library(plyr)     
library(tidyr)   
library(dplyr)

```
## R Markdown

Bring in Data

```{r readfile}

df=read.csv(file.choose(),header=TRUE)
```


#Question 1: To do data analysis identify factors that lead to attrition?
```{r factors, echo=FALSE}
control <- trainControl(method="repeatedcv", number=10, repeats=3)
# train the model
model <- train(Attrition~., data=df, method="lvq", preProcess="scale", trControl=control)
# estimate variable importance
importance <- varImp(model, scale=FALSE)
# summarize importance
print(importance)
# plot importance
plot(importance)
```
                                 Overall        Importance
Age                              3.37097096-     Yes 
BusinessTravelTravel_Frequently  3.39733632-     Yes
BusinessTravelTravel_Rarely      2.31469603-     Yes 
DepartmentResearch & Development 0.02522449-     No
DepartmentSales                  0.02332101-     No
DistanceFromHome                 3.60951514-     Yes
Education                        0.11556529-     No 
EducationFieldLife Sciences      0.05170160-     No
EducationFieldMarketing          0.53532419-     Yes
EducationFieldMedical            0.05877836-     No 
EducationFieldOther              0.30116975-     No
EducationFieldTechnical Degree   0.93258088-     Yes 
EnvironmentSatisfaction          4.16432953-     Yes
GenderMale                       1.86094740-     Yes
JobInvolvement                   3.85797696-     Yes
JobLevel                         0.27218782*     May Be
JobRoleHuman Resources           0.02915125      No 
JobRoleLaboratory Technician     2.94205866-     Yes
JobRoleManager                   0.94174152-     Yes
JobRoleManufacturing Director    0.24557120-     May Be
JobRoleResearch Director         0.06293447-     No  
JobRoleResearch Scientist        1.68219281-     Yes
JobRoleSales Executive           1.57859760-     Yes
JobRoleSales Representative      2.18420868-     Yes
JobSatisfaction                  3.59627140-     Yes
MaritalStatusMarried             1.68189531-     Yes
MaritalStatusSingle              4.92826681-     Yes
MonthlyIncome                    0.31957312-     May Be
NumCompaniesWorked               3.83038235-     Yes
OverTimeYes                      8.02574274-     Yes
PercentSalaryHike                0.48017347-     Yes
PerformanceRating                0.46448923-     May Be
RelationshipSatisfaction         2.64289543-     Yes
TotalWorkingYears                0.36628186-     May Be
TrainingTimesLastYear            1.15222224-     Yes
WorkLifeBalance                  2.87355896-     Yes
YearsAtCompany                   1.81867995-     Yes
YearsInCurrentRole               2.61057126-     Yes
YearsSinceLastPromotion          3.19963362-     Yes
YearsWithCurrManager             2.26107988-     Yes

```{r model}
str(att.df)
train <- df[1:799,]
test<- df[800:1400,]
model <- glm(Attrition ~ Age+BusinessTravel+Department+DistanceFromHome+
               Education+EducationField+EnvironmentSatisfaction+Gender+JobInvolvement+
               JobLevel+JobRole+JobSatisfaction+MaritalStatus+MonthlyIncome+NumCompaniesWorked+
               OverTime+PercentSalaryHike+PerformanceRating+RelationshipSatisfaction+
               TotalWorkingYears+TrainingTimesLastYear+WorkLifeBalance+ 
               YearsAtCompany+YearsInCurrentRole+YearsSinceLastPromotion+
               YearsWithCurrManager,family=binomial(link='logit'),data=train)
summary(model)
```

Call:
glm(formula = Attrition ~ Age + BusinessTravel + Department + 
    DistanceFromHome + Education + EducationField + EnvironmentSatisfaction + 
    Gender + JobInvolvement + JobLevel + JobRole + JobSatisfaction + 
    MaritalStatus + MonthlyIncome + NumCompaniesWorked + OverTime + 
    PercentSalaryHike + PerformanceRating + RelationshipSatisfaction + 
    TotalWorkingYears + TrainingTimesLastYear + WorkLifeBalance + 
    YearsAtCompany + YearsInCurrentRole + YearsSinceLastPromotion + 
    YearsWithCurrManager, family = binomial(link = "logit"), 
    data = train)

Deviance Residuals: 
     Min        1Q    Median        3Q       Max  
-1.75508  -0.43147  -0.20751  -0.06315   2.95812  

Coefficients:
                                   Estimate Std. Error z value Pr(>|z|)    
(Intercept)                      -1.148e+01  8.839e+02  -0.013 0.989636    
Age                               3.477e-04  5.885e-04   0.591 0.554632    
BusinessTravelTravel_Frequently   1.645e+00  5.441e-01   3.023 0.002503 ** 
BusinessTravelTravel_Rarely       9.179e-01  4.832e-01   1.900 0.057485 .  
DepartmentResearch & Development  1.402e+01  8.839e+02   0.016 0.987341    
DepartmentSales                   1.391e+01  8.839e+02   0.016 0.987444    
DistanceFromHome                  4.605e-02  1.539e-02   2.992 0.002773 ** 
Education                        -1.315e-01  1.259e-01  -1.045 0.296176    
EducationFieldLife Sciences      -1.384e+00  1.217e+00  -1.137 0.255545    
EducationFieldMarketing          -1.552e+00  1.292e+00  -1.201 0.229743    
EducationFieldMedical            -1.458e+00  1.213e+00  -1.201 0.229594    
EducationFieldOther              -9.641e-01  1.297e+00  -0.743 0.457240    
EducationFieldTechnical Degree   -8.324e-01  1.257e+00  -0.662 0.507777    
EnvironmentSatisfaction          -3.202e-01  1.201e-01  -2.666 0.007665 ** 
GenderMale                        1.358e-01  2.661e-01   0.510 0.609955    
JobInvolvement                   -8.595e-01  1.807e-01  -4.756 1.97e-06 ***
JobLevel                         -2.407e-01  4.663e-01  -0.516 0.605703    
JobRoleHuman Resources            1.517e+01  8.839e+02   0.017 0.986306    
JobRoleLaboratory Technician      1.142e+00  6.388e-01   1.787 0.073907 .  
JobRoleManager                    8.953e-02  1.210e+00   0.074 0.941010    
JobRoleManufacturing Director    -1.272e+00  9.129e-01  -1.393 0.163550    
JobRoleResearch Director         -2.302e+00  1.488e+00  -1.547 0.121891    
JobRoleResearch Scientist         5.533e-01  6.409e-01   0.863 0.387954    
JobRoleSales Executive            8.606e-01  1.388e+00   0.620 0.535338    
JobRoleSales Representative       2.504e+00  1.468e+00   1.706 0.088074 .  
JobSatisfaction                  -5.917e-01  1.196e-01  -4.946 7.56e-07 ***
MaritalStatusMarried              9.688e-01  4.067e-01   2.382 0.017223 *  
MaritalStatusSingle               1.828e+00  4.107e-01   4.451 8.57e-06 ***
MonthlyIncome                     1.316e-04  1.266e-04   1.039 0.298598    
NumCompaniesWorked                2.530e-01  5.936e-02   4.262 2.03e-05 ***
OverTimeYes                       2.027e+00  2.738e-01   7.403 1.33e-13 ***
PercentSalaryHike                 8.959e-03  5.513e-02   0.163 0.870904    
PerformanceRating                 1.893e-01  5.697e-01   0.332 0.739715    
RelationshipSatisfaction         -2.000e-01  1.171e-01  -1.709 0.087498 .  
TotalWorkingYears                -1.256e-01  3.913e-02  -3.210 0.001326 ** 
TrainingTimesLastYear            -2.892e-01  1.133e-01  -2.553 0.010687 *  
WorkLifeBalance                  -6.008e-01  1.748e-01  -3.437 0.000588 ***
YearsAtCompany                    1.127e-01  5.540e-02   2.034 0.041967 *  
YearsInCurrentRole               -1.649e-01  6.497e-02  -2.538 0.011147 *  
YearsSinceLastPromotion           2.649e-01  6.163e-02   4.298 1.72e-05 ***
YearsWithCurrManager             -1.333e-01  6.955e-02  -1.916 0.055363  
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

(Dispersion parameter for binomial family taken to be 1)

    Null deviance: 706.43  on 798  degrees of freedom
Residual deviance: 425.69  on 758  degrees of freedom
AIC: 507.69

Number of Fisher Scoring iterations: 15


#Question 2   Top three factors that contribute to turnover (backed up by evidence provided by analysis)
##Top Three (Positive) factors
OverTimeYes                       2.027e+00  2.738e-01   7.403 1.33e-13 ***
MaritalStatusSingle               1.828e+00  4.107e-01   4.451 8.57e-06 ***
YearsSinceLastPromotion           2.649e-01  6.163e-02   4.298 1.72e-05 ***
##Top Three (Negative) factors
JobSatisfaction                  -5.917e-01  1.196e-01  -4.946 7.56e-07 ***
JobInvolvement                   -8.595e-01  1.807e-01  -4.756 1.97e-06 ***
WorkLifeBalance                  -6.008e-01  1.748e-01  -3.437 0.000588 **
#Top Three factors
OverTimeYes                      8.02574274-     Yes
MaritalStatusSingle              4.92826681-     Yes
EnvironmentSatisfaction          4.16432953-     Yes
JobInvolvement                   3.85797696-     Yes
NumCompaniesWorked               3.83038235-     Yes
DistanceFromHome                 3.60951514-     Yes
JobSatisfaction                  3.59627140-     Yes
BusinessTravelTravel_Frequently  3.39733632-     Yes
Age                              3.37097096-     Yes
YearsSinceLastPromotion          3.19963362-     Yes

```{r packages, echo=FALSE}
#Get the details of each explanatory variable and distribution 
install.packages("psych") ## install
library(psych)
describe(att.df)
table(Attrition)
table(BusinessTravel)
table(Department)
table(EducationField)
table(Education)
table(EnvironmentSatisfaction)
table(Gender)
table(JobInvolvement)
table(JobRole)
table(JobLevel)
table(JobSatisfaction)
table(MaritalStatus)
table(OverTime)
```
# Explanatory Variable Distribution
table(Attrition)
Attrition
  No  Yes 
1228  242 
#Dataset with only 242 left employees, 1228 current employees 
table(BusinessTravel)
BusinessTravel
       Non-Travel Travel_Frequently     Travel_Rarely 
              142               282              1046 
#Dataset with only 282 frequent travellers
table(Department)
Department
       Human Resources Research & Development                  Sales 
                    63                    961                    446 
#Dataset has 446 Sales Team
table(EducationField)
EducationField
 Human Resources    Life Sciences        Marketing          Medical            Other Technical Degree 
              27              606              159              464               82              132 
table(Education)
Education
  1   2   3   4   5 
170 282 572 398  48 
table(EnvironmentSatisfaction)
EnvironmentSatisfaction
  1   2   3   4 
284 287 453 446 
#Dataset with only 284 high satisfaction
table(Gender)
Gender
Female   Male 
   588    882 
#Dataset has majority Male Employees,40 percent females ,60 percent males.
table(JobInvolvement)
JobInvolvement
  1   2   3   4 
 83 375 868 144 
table(JobInvolvement)
#Dataset has only 144 employees have high job involvement.
table(JobRole)
JobRole
Healthcare Representative           Human Resources     Laboratory Technician                   Manager 
                      131                        52                       259                       102 
   Manufacturing Director         Research Director        Research Scientist           Sales Executive 
                      145                        80                       292                       326 
     Sales Representative 
                       83 
table(JobLevel)
JobLevel
  1   2   3   4   5 
543 534 218 106  69 
table(JobSatisfaction)
JobSatisfaction
  1   2   3   4 
289 280 442 459 
#Dataset has only 459 employees have high job satisfaction, 289 have very low job satisfaction.
table(MaritalStatus)
MaritalStatus
Divorced  Married   Single 
     327      673      470 
#Dataset has only 470 employees are single, they are higher chance leaving company.
table(OverTime)
OverTime
  No  Yes 
1054  416 
#Dataset has 1054 employees have no overtime, so they are candidates, they have high chance of leaving company.

#Invidual Explanatory Variables Relationship with Attrition
```
mytable<- xtabs(~PerformanceRating+Attrition,data=att.df)
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~OverTime+Attrition,data=att.df)
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~WorkLifeBalance+Attrition,data=att.df) 
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~JobRole+Attrition,data=att.df) 
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~NumCompaniesWorked+Attrition,data=att.df)
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~MaritalStatus+Attrition,data=att.df) 
round(prop.table(mytable,1)*100,2) 
mytable<- xtabs(~Gender+Attrition,data=att.df) 
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~EnvironmentSatisfaction+Attrition,data=att.df) 
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~BusinessTravel+Attrition,data=att.df) 
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~EducationField+Attrition,data=att.df) 
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~JobSatisfaction+Attrition,data=att.df)
round(prop.table(mytable,1)*100,2)
mytable<- xtabs(~Education+Attrition,data=att.df) 
round(prop.table(mytable,1)*100,2)
```
## Including Explanatory Variable Relationship with Attrition
> mytable<- xtabs(~PerformanceRating+Attrition,data=att.df)
> round(prop.table(mytable,1)*100,2)
                 Attrition
PerformanceRating    No   Yes
                3 83.60 16.40
                4 83.19 16.81
#High Performance Rating less chance of leaving company.
> mytable<- xtabs(~OverTime+Attrition,data=att.df)
> round(prop.table(mytable,1)*100,2)
        Attrition
OverTime    No   Yes
     No  88.43 11.57
     Yes 71.15 28.85
#No Overtime high chance of leaving company.
> mytable<- xtabs(~WorkLifeBalance+Attrition,data=att.df) 
> round(prop.table(mytable,1)*100,2)
               Attrition
WorkLifeBalance    No   Yes
              1 72.50 27.50
              2 84.01 15.99
              3 84.43 15.57
              4 83.01 16.99
#Low WorkLife Balance high chance of leaving company.
> mytable<- xtabs(~JobRole+Attrition,data=att.df) 
> round(prop.table(mytable,1)*100,2)
                           Attrition
JobRole                        No   Yes
  Healthcare Representative 87.79 12.21
  Human Resources           78.85 21.15
  Laboratory Technician     79.92 20.08
  Manager                   92.16  7.84
  Manufacturing Director    89.66 10.34
  Research Director         96.25  3.75
  Research Scientist        81.16 18.84
  Sales Executive           84.05 15.95
  Sales Representative      63.86 36.14
#Sales Representative Job Role has high chance of leaving company.
> mytable<- xtabs(~NumCompaniesWorked+Attrition,data=att.df)
> round(prop.table(mytable,1)*100,2)
                  Attrition
NumCompaniesWorked    No   Yes
                 0 88.32 11.68
                 1 82.92 17.08
                 2 85.62 14.38
                 3 87.42 12.58
                 4 87.77 12.23
                 5 77.78 22.22
                 6 71.43 28.57
                 7 77.03 22.97
                 8 85.71 14.29
                 9 73.08 26.92
#Employee worked more companies high chance of leaving company.
> mytable<- xtabs(~MaritalStatus+Attrition,data=att.df) 
> round(prop.table(mytable,1)*100,2) 
             Attrition
MaritalStatus    No   Yes
     Divorced 89.60 10.40
     Married  86.03 13.97
     Single   75.74 24.26
#Singles-Not Married high chance of leaving company.
> mytable<- xtabs(~Gender+Attrition,data=att.df) 
> round(prop.table(mytable,1)*100,2)
        Attrition
Gender      No   Yes
  Female 85.20 14.80
  Male   82.43 17.57
#Males are little higher chance of leaving company compare to females.
> mytable<- xtabs(~EnvironmentSatisfaction+Attrition,data=att.df) 
> round(prop.table(mytable,1)*100,2)
                       Attrition
EnvironmentSatisfaction    No   Yes
                      1 74.30 25.70
                      2 86.76 13.24
                      3 85.43 14.57
                      4 85.43 14.57
#Low EmploymentSatisfaction has high chance of leaving company.
> mytable<- xtabs(~BusinessTravel+Attrition,data=att.df) 
> round(prop.table(mytable,1)*100,2)
                   Attrition
BusinessTravel         No   Yes
  Non-Travel        92.25  7.75
  Travel_Frequently 74.11 25.89
  Travel_Rarely     84.89 15.11
#Employees who travel frequently have high chance of leaving company.
> mytable<- xtabs(~EducationField+Attrition,data=att.df) 
> round(prop.table(mytable,1)*100,2)
                  Attrition
EducationField        No   Yes
  Human Resources  74.07 25.93
  Life Sciences    84.98 15.02
  Marketing        82.39 17.61
  Medical          83.84 16.16
  Other            84.15 15.85
  Technical Degree 78.79 21.21
> mytable<- xtabs(~JobSatisfaction+Attrition,data=att.df)
> round(prop.table(mytable,1)*100,2)
               Attrition
JobSatisfaction    No   Yes
              1 77.85 22.15
              2 82.14 17.86
              3 83.71 16.29
              4 87.80 12.20
#Low JobSatisfaction has high chance of leaving company.
> mytable<- xtabs(~Education+Attrition,data=att.df) 
> round(prop.table(mytable,1)*100,2)
         Attrition
Education    No   Yes
        1 82.35 17.65
        2 82.27 17.73
        3 83.22 16.78
        4 84.92 15.08
        5 87.50 12.50

#Education has not having much influence of Attriton/leaving company.

#Question 3: Job role specific trends that may exist in the data set.

```{r role}
```
##Job Role Specific Trends
                          Attrition
JobRole                        No   Yes
  Healthcare Representative 87.79 12.21
  Human Resources           78.85 21.15
  Laboratory Technician     79.92 20.08
  Manager                   92.16  7.84
  Manufacturing Director    89.66 10.34
  Research Director         96.25  3.75
  Research Scientist        81.16 18.84
  Sales Executive           84.05 15.95
  Sales Representative      63.86 36.14
#Sales Representative Job Role has high chance of leaving company.
#Human Resources Job Role next high chance of leaving company.
#Research Scientist Job Role has next high chance of leaving company.

```
#Question 4: Any other interesting trends and observations from your analysis.
    ##The dataset is imbalanced with the only 242 left employees-18 percent of population, 1228 current employees-82 percent of population.
	##People who have to work overtime show higher percentage of leavers compared to their counterparts.
	##Single employees show the highest percentage of leavers, compared to Married and Divorced counterparts.
	##The strongest positive correlations with the target features are: Performance Rating, Monthly Rate, Num Companies Worked, Distance From Home.
    ##The strongest negative correlations with the target features are: Total Working Years, Job Level, Years In Current Role, and Monthly Income.
    ##About 10% of leavers left when they reach their 2-year anniversary at the company.
    ##People who live further away from their work show higher percentage of leavers compared to their counterparts.
    ##People who travel frequently show higher percentage of leavers compared to their counterparts.
    ##Employees that have already worked at many companies previously,higher percentage of leavers compared to their counterparts.

```{r extra}

#Anyother extra details


```










